\LARGE \bf \pyRobots{}, a toolset for robot executive control

### 2014-09-23 paper.tex
Implemented Ayberk's review. Thanks a lot!

Presented is \pyRobots{}, a new software toolkit for the executive control of
complex robotic systems. Borrowing ideas from previous tools like {\sc
urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
develop (pre-emptive) concurrent and event-based executive controllers.

Designed in a \emph{bottom-up} fashion, out of the need for a practical,
unobstrusive tool suitable for rapid prototyping of complex interaction
scenarios, \pyRobots{} also exposes several simple yet convenient abstractions for
physical resources management and pose representation.

Finally, and while middleware-agnostic, it features advanced integration with
the \textit{de-facto} standard ROS middleware.

@@

While individual decision-making components (like task planners) are studied in
(relative) isolation since years, the \emph{orchestration} issue, with questions
like \textit{When to start them? Which one should be selected? How to react to a
new situation in a timely manner?} remains difficult to address in a generic
way.

Many approaches have been devised (we review them in the next section) that
include \emph{Finite State Machines} (FSM), domain-specific languages (in
particular, logic languages) or agent-based frameworks.

These tools adopt principled approaches, often with solid theoretical
contributions. However, it seems that none of them gained broad acceptance in
the robotic community, and many of us resort to write ad-hoc scripts, suitable
for a single application/demonstration/experiment and neither reliable nor
extensible.

We hypothesize that the software architecture design that these tools enforce,
combined with their general lack of practicality (unfamiliar language; lack of
bindings for a given robot or middleware; difficulty to install or setup;
non-trivial deployment) explain this low level of acceptance more than any
particular intrinsic weaknesses.

This article introduces our attempt at designing an unobtrusive execution
control toolkit, that aims at addressing the \emph{practicality} issue: it has
been designed and implemented from the bottom-up, starting from actual needs
when running large robotic systems (including Willow Garage's PR2 and
Aldebaran's Nao) in dynamic and partially unpredictable environments (typically,
loosely constrained human-robot interaction scenarios).

Instead of an \emph{environment} or a \emph{framework}, which would suggest
rigid design and development guidelines, \pyRobots{} can be conceived as a set
of software helpers to write parallel, event-based, high-level robot
controllers.  It is written for and in Python, which ensure familiarity, fast
development cycles and broad support for interfacing with existing robots and
middlewares. \pyRobots{} is indeed \emph{not} a middleware: it purposefully
provides no mechanisms to connect to other modules or components. It instead
\emph{relies} on one (or several) middleware to actually communicate with the
robot and perform actions.

Also, \pyRobots{} does not provide formal models or verifiability: in its
current state, it does not attempt to contribute in this field. It focuses
instead on providing a lightweight tool to control robots in environments that
require carrying out many tasks in parallel, in an unpredictable, event-based
manner. As such, \pyRobots{} is focusing on use-cases ranging from prototyping
to mid-size experiments.

@@

Any robotic experiment requires a certain amount of high-level control to
supervise the robot behaviour, and, expectedly, many projects have attempted to
design and provide tools for this task. Degroote provides an in-depth review of
the existing techniques and tools in~\cite{degroote2012architecture}. We focus
this review on tools dedicated to the explicit implementation of high-level
behaviours, excluding complete control stacks (that would typically include a
middleware and hardware abstractions), meta-tools (like component generators or
model checkers) and (semi-)automatic techniques (like emerging behaviours or
learning by demonstration).  In contrast with middlewares for which the robotic
community relies on a few \emph{de-facto} standards (ROS, YARP, OpenRTM), we
want to underline that no such widely-used softwares exist for execution
control, and we believe that most of today's robot behaviours are implemented as
\emph{ad-hoc} scripts.

We identify three categories of execution control tools: \emph{finite state
machines}, \emph{agent-based} controllers and
\emph{domain-specific languages} (DSL), including \emph{embedded domain-specific
languages} (eDSL,~\cite{joyeux2011robot}). These categories overlap to a certain
extend, but they should reasonably reflect the motivational concepts/ideas that
gave birth to the specific tools.

\paragraph{Finite State Machines} (or FSM) are the most well-known paradigm to
describe how a robot goes from one \emph{state} (typically, performing an
action) to another state when some condition becomes true or some event is
triggered. Common FSM implementations in robotics include {\sc
rfSM}~\cite{klotzbucher2010orocos} (part of the OROCOS framework) and
{\sc smach}~\cite{bohren2010smach}. State machine semantics are well
understood~\cite{harel1996statemate} and provide a well structured model,
suitable for meta-reasoning (like verification). However, FSM are also
acknowledged to be ill-suited for complex systems (where the number of state
easily ``explodes'') or relatively dynamic and unpredictable environments where
it can prove difficult to \textit{a priori} exhaustively list transitions.

@@

The T-REX architecture~\cite{mcgann2007trex} partitions the execution control
problem into several agents, each one composed of a (temporal) planner and an
execution layer. At run-time, the different agents are constantly synchronised
to maintain the consistency of the global plan. ROAR~\cite{degroote2011roar}
augments this idea with a logic mechanism allowing the system to automatically
select and start agents depending on the task at hand.

@@

Domain-specific languages (DSL), in the broadest sense, represent the largest
body of literature on high-level behaviour programming. They can either be
actual domain-specific languages like PRS~\cite{ingrand1996prs},
RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently, {\sc
urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
languages like Python or Ruby, or logic languages like Prolog) to provide
high-level paradigms and control structures well suited to the programming of
robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
roby}~\cite{joyeux2009plan} (based on Ruby), {\sc teer}~\cite{magnenat2012teer}
(based on Python), are examples of such eDSL.

\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a
similar fashion to {\sc roby} or {\sc teer}), that focuses on high-level
behaviours implementation. It seeks unobtrusive interaction with the other
components of the robot software stack, using Python as ubiquitous binding
language. Similar to {\sc urbi}, it focuses on transparent pre-emptive
concurrency and event-based programming.  Also, \pyRobots{}' model favours
hierarchical task composition over plans, and does not provide by itself any
task planning service, delegating this to optional external tools.

@@

This section introduces the main concepts of \pyRobots{},
and provide an overview of the features and mechanisms. Since \pyRobots{} is an
open-source software, the interested reader can also refer to the source code
and documentation, available from~\url{https://github.com/chili-epfl/pyrobots}.

@@

\pyRobots{} core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware like ROS) and the \textbf{state} of the robot
(typically a set proxies to the sensors, or a connector to a database, etc.,
accessed through a key-value container -- a dictionary-like object).

@@

The user can list \textbf{events} that the \textbf{robot} must monitor, and
attach callback to them.

Finally, the user can declare arbitrary \textbf{resources}, usually
corresponding to the robot physical components (for instance, the wheels, the
camera) and guarantee exclusive access to the required resource to specific
actions.

@@

\emph{Actions} in \pyRobots{} are asynchronous by default. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
function annotated with the \python|@action| decorator, and are invoked
like any other function.  Since they are asynchronous, these functions may
perfectly block or even never return (typically useful to implement continuous
background behaviours).

Actions can also call each other (nested actions), allowing natural
decomposition of complex tasks into sub-actions and atomic actions (we note
hereafter $depth$ the nesting level of an action, \ie the maximum number of
calls to sub-actions until an atomic action is reached).

\emph{Signaling:} Behaviours often need to be paused or cancelled.  This raises
an issue that specifically needs to be addressed in robotics: before suspending
an action, one often wants to first bring back the robot into some form of rest
state (\eg one would not want to suddenly cancel a \python|walk| action in the
middle of a step without first bringing back the foot to the ground). \pyRobots{}
addresses this issue through signals, and extends standard Python thread into
\emph{signaling threads}, relying on Python's usual syntax for exception
handling. Listing~\ref{lst:signals} illustrates this mechanism. The action is
instantiated and started at line 8, and cancelled at line 10. This send a
\python|ActionCancelled| signal to the action that the user can handle using an
\python|except| statement.

@@

    action = robot.safe_walk()
    time.sleep(1)
    action.cancel()
\end{pythoncode}
\caption{\textbf{Handling a cancellation signal} After one second, the
\python|safe_walk| action is cancelled. This sends the signal
\python|ActionCancelled| to the action, that can appropriately handle it inside
the \python|except| block.}
\label{lst:signals}
\end{listing}

Pausing actions is more complex (the action internal state must be saved, a mean
to resume the action where it was interrupted must be provided, the user must be
given ways to code transition behaviours when pausing and resuming) and is not
currently supported by \pyRobots{}.

@@

Sharing the same syntax with {\sc urbi}, events in \pyRobots{} can by monitored
through the methods \python|on| (one shot event) and \python|whenever|
(continuous event monitoring). Event conditions can be either a predicate (that
must take as single argument the robot instance) or a pair of $\{\text{state
key}, \text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$.
Examples of event usage can be seen at lines 33--38 of
listing~\ref{croquignole_no_move}.

@@

\subsection{Resources management}

Physical robot resources (mainly actuators, but possibly sensors as
well) usually mandate exclusive use (it is seldom desirable that two different actions
control the wheels of a robot at the same time). \pyRobots{} provides a resource
locking mechanism that relies on the \python|@lock| annotation: an action
annotated with \python|@lock(ARMS)| (Listing~\ref{lst:resources}, l.6) will
acquire (or wait for) the \python|ARM| resource when being executed.

This mechanism is generic and \pyRobots{} makes no assumptions concerning the
actual resources available/meaningful to a specific robot. The
programmer define them himself, either as stand-alone or as compound
resources (listing~\ref{lst:resources}, l.1--3).

A given action can lock several distinct resources (with the corresponding
number of \python|@lock| annotations. When executed, the
action acquires a lock on the resource (and all dependent resources in the case
of a compound resource) which is released upon action completion. If the lock is
already acquired, two strategies are possible: either wait or \emph{fail fast}
(\python|@lock(..., wait=False)|, l.16).

@@

Managing poses, transformations, frames in real-world scenarios is not an easy
task, mainly because it seems that every single library, middleware or robotic
platform relies on its own conventions (rotation representations, in particular,
are a common source of inconsistencies -- quaternions, (the many) Euler
conventions, rotation matrices, Rodrigues, yaw-pitch-roll notations being all
rather frequent). Within the ROS ecosystem, the Transform Library ({\sc
tf},~\cite{foote2013tf}) has played a major role to make it convenient to work
with frames, and poses can always be expressed and shared in the most convenient
local frame, {\sc tf} taking care of required intermediate transformation.

\pyRobots{} pose management builds on this idea to offer flexible pose
representations \emph{within} the control scripts. Everytime a 6D pose needs to
be specified (like a navigation goal or a target pose for an arm), it can be
expressed as a simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt
map} is then assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y,
z, rx, ry, rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx,
ry, rz, frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x,
y, z, qx, qy, qz, qw, frame\}$ or by a dictionary that defines any subset of
these 8 values.  Internally, \pyRobots{} normalizes and stores the poses in
their full 8-tuple expansion, as illustrated in listing~\ref{lst:poses}.

@@

Using helper methods like \python|inframe(<pose>, <other frame>)| or
\python|pantilt(<pose>, <ref frame>)|, the programmer can then
transform a given pose into a different frame or different representation.

Required transformations are computed by \pyRobots{}, and the programmer can
provide new ``sources'' of frame transformations through the instanciation of
\python|FrameProviders| that takes advantage of robot- and middleware-specific
infrastructures (for example, ROS {\sc tf}, Aldebaran's {\tt NAOqi} frame
conventions, etc.).

@@

Running robots in real-world, \ie dynamic and unpredictable, environments
entails repetitive coding-testing-debugging development cycles. \pyRobots{}
provides several tools to support this development phase. A special attention
was paid to the logging infrastructure. It relies on the versatile Python
built-in logging tools, which support configurable parallel logging streams
(typically, a full debug stream is logged to a file while a less verbose stream
is output on the console at run-time).

\pyRobots{} logs are designed to allow full \emph{post-mortem} replay, and we
provide a graphical interface to display the actions and resource usage over
time (Figure~\ref{log_view}).

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{log}
        \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
        experiment presented at section~\ref{croq}}
        \label{log_view}
\end{figure}

Debugging and introspection support is provided at run-time through a range of
off-the-shell Python debuggers. Remote debugging is also supported with standard
tools like {\sc rpdb2}~\footnote{\url{http://winpdb.org}}. Thanks to the dynamic
nature of Python, these debuggers not only provide insights on failures, but
also provide full introspection and run-time manipulation of the application
state.

@@

The development of \pyRobots{} started in 2011, driven by the requirements of a
18min long theatre performance between a human actor and a robot. Since that
time, the toolkit has iteratively matured, and has been deployed and used on
several projects and platforms. We first briefly recap these deployments with
references to the corresponding publications, and then present our latest study,
in a kindergarten, that recently acted as a stress-test for the library.

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
        \toprule
        Study                 & Robot & Middlewares \\ \midrule
        Roboscopie            & PR2       & ROS, {\sc GenoM} \\
        Interactive Grounding & PR2, LAAS-CNRS's Jido & ROS, {\sc GenoM} \\
        CoWriter              & Nao       & ROS, NAOqi  \\
        Ranger at kindergarten & EPFL's Ranger & ROS, aseba  \\
        \bottomrule
    \end{tabular}
    \caption{Overview of significant experiments and studies where \pyRobots{}
    has been used.}
\end{table}

\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
a theatre performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}.
The toolkit was created out of the need to control a complex robot (a Willow
Garage PR2) in a structured way, as required for a live performance with a
human\footnote{The original \pyRobots{} script used for the performance can be
accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
not reflect anymore the current toolkit usage.}

A total of 47 \pyRobots{} actions were written for this event
(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolkit
proved to be especially convenient to interleave calls to both ROS-based and
{\sc GenoM}-based~\cite{mallet2010genom3} components through their respective
Python bindings.

\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
script language, essential for the robot to ``practise'' and adapt to the
vision of the director.

@@

\subsubsection{Interactive Grounding} This study on perspective-taking and grounding
in human-robot interaction~\cite{lemaignan2013talking} is one of the several
experiments that was built on top of the LAAS-CNRS cognitive
architecture~\cite{lemaignan2014human}. In this large architecture, \pyRobots{}
is one of the two available executive controllers, and integrates with a task
planner, a symbolic knowledge base, several motion planner and a natural
language processor. \pyRobots{} was favoured over the second execution
controller (based on PRS) for smaller studies, as a more powerful alternative to
bare \textit{ad-hoc} scripts.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{cowriter}
        \caption{Children teaching an Aldebaran Nao interacting how to
        handwrite.}
        \label{expe-cowriter}
\end{figure}

\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
(Figure~\ref{expe-cowriter}) involves an Aldebaran Nao robot and young students
in a \emph{learning by teaching} interaction: the children have to teach the
robot how to handwrite. This experiment involved a dual ROS/NAOqi environment,
bridged within \pyRobots{}. The high-level behaviour was implemented as a finite
state machine ({\sc SMACH}) whose state corresponded to \pyRobots{} actions.
While the execution controller was subsequently simplified (towards a tighter
interaction scenario), this illustrates how \pyRobots{}, relying on the
ubiquitous Python language, can possibly blend into and interact with existing
tools.

@@

We organized in June 2014 a pilot study in a kindergarten (on campus) to
stress-test the hardware and software of the Ranger
robot~\cite{mondada2014ranger}. The Ranger robot is an autonomous robot designed
for interaction with children. As it was primarily designed to support children
with tidying up, it is shaped as a wooden ``box on wheels''
(Figure~\ref{expe-nursery}). Hardware-wise, it features an RGB-D camera, IR
sensors, a range and bearing sensor, a physical front bumper, a scale, a
removable pacifier (used as a soft on/off switch by the children) and it is
covered with capacitive tactile panels. The wooden panels hide 186 RGB LEDs,
enabling various light patterns, it can play sounds, the two eyes can be fully
controlled (2 DoFs pupils and actuated eye-lids). Wheel controllers are designed
such as they switch to freewheel as soon as the robot is pushed by the children.
The robot is powered by an embedded Gumstix Overo board (ARM7l) running a custom
Linux image, assisted by three custom microcontroller boards (for low-level
hardware interfacing).

Software-wise, all the user-facing behaviours are implemented with \pyRobots{},
and the communication with the hardware is abstracted through a mix of ROS
and Aseba~\cite{magnenat2011aseba} nodes.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviours of the robot are implemented via \pyRobots{}.}
        \label{expe-nursery}
\end{figure}

>>>>

Presented is \pyRobots{}, a new open-source software toolkit for the executive
control of complex robotic systems. Borrowing ideas from previous tools like
{\sc urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
develop (pre-emptive) concurrent and event-based executive controllers.

Designed in a \emph{bottom-up} fashion, out of the need for a practical,
unobstrusive tool suitable for rapid prototyping of complex interaction
scenarios, \pyRobots{} also exposes several simple yet convenient abstractions
for physical resources management and pose representation. While
middleware-agnostic, it features specific integration with the ROS middleware.

Experimental deployments and stress-tests on several robotic platforms
(including PR2 and Nao) in dynamic human environments are reported.

@@

While individual decision-making components (such as task planners) have been
studied in (relative) isolation for decades, the \emph{orchestration} issue,
with questions such as \textit{When to start them? Which one should be selected?
How to react to a new situation in a timely manner?} remains difficult to
address in a generic way.

Many approaches have been devised that
include \emph{Finite State Machines} (FSM), domain-specific languages (in
particular, logic languages) or agent-based frameworks; these are reviewed in
the next section.

These tools adopt principled approaches, often with solid theoretical
contributions. However, it seems that none of them gained broad acceptance in
the robotic community, and many of us resort to write ad-hoc scripts, suitable
for a single application/demonstration/experiment and neither reliable nor
extensible. We hypothesize that the software architecture that these tools
enforce, combined with their general lack of practicality (unfamiliar language;
lack of bindings for a given robot or middleware; difficulty to install or
setup; non-trivial deployment) explain this low level of acceptance more than
any particular intrinsic weaknesses.

This article introduces our attempt at designing an unobtrusive execution
control toolset that aims at addressing the \emph{practicality} issue: it has
been designed and implemented from the bottom-up, starting from actual needs
when running large robotic systems (including robots such as PR2 or Nao) in
dynamic and partially unpredictable environments, such as loosely constrained
human-robot interaction scenarios.

Instead of an \emph{environment} or a \emph{framework}, which would suggest
rigid design and development guidelines, \pyRobots{} can be conceived as a set
of software libraries to write parallel, event-based, high-level robot
controllers.  It is written for and in Python, which ensure familiarity, fast
development cycles and broad support for interfacing with existing robots and
middlewares. \pyRobots{} is indeed \emph{not} a middleware: it purposefully
provides no mechanisms to connect to other modules or components. It instead
\emph{relies} on one (or several) middleware(s) to actually communicate with the
robot.

@@

Any robotic experiment requires a certain amount of high-level control to
supervise the robot behaviour, and, expectedly, many projects have attempted to
design and provide tools for this task. Degroote provides an in-depth review of
the existing techniques and tools in~\cite{degroote2012architecture}. We focus
this review on tools dedicated to the explicit implementation of high-level
behaviours, excluding complete control stacks (that would typically include a
middleware and hardware abstractions), meta-tools (such as component generators
or model checkers) and (semi-)automatic techniques (such as emerging behaviours
or learning by demonstration). Unlike \emph{middlewares} where a few tools --
ROS, YARP, {\sc orocos}-rtt, OpenRTM -- are the \emph{de facto} standards, no tool really
dominate the field of execution control, and we believe instead that most of
today's robot behaviours are implemented as \emph{ad-hoc} scripts.

We identify three categories of execution control tools: \emph{finite state
machines}, \emph{agent-based} controllers and
\emph{Domain-Specific Languages} (DSL), including \emph{embedded Domain-Specific
Languages} (eDSL,~\cite{joyeux2011robot}). These categories overlap to a certain
extent, but still reflect reasonably well the motivational concepts/ideas that
gave birth to the specific tools.

\paragraph{Finite State Machines} They are the most well-known paradigm to
describe how a robot goes from one \emph{state} (typically, performing an
action) to another state when some condition is satisfied or some event is
triggered. Common FSM implementations in robotics include {\sc
rfSM}~\cite{klotzbucher2010orocos} (part of the {\sc orocos} framework) and
{\sc smach}~\cite{bohren2010smach}. State machine semantics are well
understood~\cite{harel1996statemate} and provide a well structured model,
suitable for meta-reasoning (such as verification). However, FSM are also
acknowledged to be ill-suited for complex systems (where the number of state
easily ``explodes'') or relatively dynamic and unpredictable environments where
it can prove difficult to exhaustively list transitions \textit{a priori}.

@@

The T-REX architecture~\cite{mcgann2007trex} partitions the execution control
problem into several agents, each one composed of a (temporal) planner and an
execution layer. At run-time, the different agents are constantly synchronised
to maintain the consistency of the global plan. ROAR~\cite{degroote2011roar}
augments this idea with a logic mechanism, allowing the system to automatically
select and start agents depending on the task at hand.

@@

In the broadest sense, they represent the largest body of literature on
high-level behaviour programming. They can either be actual domain-specific
languages such as PRS~\cite{ingrand1996prs}, RPL~\cite{mcdermott1993reactive},
Golog~\cite{levesque1997golog} (and its derivatives such as {\sc
ReadyLOG}~\cite{ferrein2007robot}) or, more recently, {\sc
urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
DSL~\cite{joyeux2011robot}. eDSL are extensions of existing languages (typically, dynamic
languages such as Python or Ruby, or logic languages such as Prolog) to provide
high-level paradigms and control structures well suited to the programming of
robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
roby}~\cite{joyeux2009plan} (based on Ruby), {\sc teer}~\cite{magnenat2012teer}
(based on Python), are examples of such eDSL.

\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a
similar fashion to {\sc roby} or {\sc teer}), that focuses on high-level
behaviour implementation. Unlike larger stacks, \pyRobots{} does not provide
formal models or verifiability. It aims to provide unobtrusive interaction with
the other components of the robot software stack, using Python as the ubiquitous
binding language. Similar to {\sc urbi}, it focuses on transparent pre-emptive
concurrency and event-based programming.  Also, \pyRobots{}' model favours
hierarchical task composition over plans, and does not provide by itself any
task planning service.

@@

This section introduces the main concepts of \pyRobots{}, and provide an
overview of the features and mechanisms. The interested reader can refer to
the source code and documentation, available
from~\url{https://github.com/chili-epfl/pyrobots}.

@@

\pyRobots{} core concepts are simple: on one hand, the user creates a
\textbf{robot} as an instance of a \python|GenericRobot| that encapsulates the
different low-level \textbf{controllers} (proxies to actuators, typically
provided by a middleware such as ROS) and the \textbf{state} of the robot
(typically a set of proxies to the sensors accessed through a key-value
container).

@@

The user can list \textbf{events} that the \textbf{robot} must monitor, and
attach callbacks to them.

Finally, the user can declare arbitrary \textbf{resources}, usually
corresponding to the robot's physical components (for instance, the wheels, the
camera) and guarantee exclusive access to these resource to a specific set of
actions.

@@

\emph{Actions} in \pyRobots{} are asynchronous by default. They are implemented as
\emph{futures} (also known as \emph{promises}) and therefore executed in
independent threads. To the developer, however, they appear as regular Python
functions annotated with the \python|@action| decorator, and are invoked
like any other function.  Since they are asynchronous, these functions may
block or even never return (typically useful for implementing continuous
background behaviours).

Actions can also call each other (nested actions), allowing natural
decomposition of complex tasks into sub-actions (the nesting depth of an action
is hereafter noted $depth$).

Behaviours often need to be paused or cancelled.  This raises
an issue that specifically needs to be addressed in robotics: before suspending
an action, one often wants to first bring back the robot into some form of rest
state (\eg one would not want to suddenly cancel a \python|walk| action in the
middle of a step without first bringing back the foot to the ground). \pyRobots{}
addresses this issue through signals, and extends the standard Python thread into
\emph{signaling threads}, relying on Python's usual syntax for exception
handling. Listing~\ref{lst:signals} illustrates this mechanism. The action is
instantiated and started at line 8, and cancelled at line 10. This sends an
\python|ActionCancelled| signal to the action that the user can handle using an
\python|except| statement.

@@

    action = robot.safe_walk()
    time.sleep(1)
    action.cancel()
\end{pythoncode}
\caption{\textbf{Handling a cancellation signal} After one second, the
\python|safe_walk| action is cancelled. This sends the signal
\python|ActionCancelled| to the action that can appropriately handle it inside
the \python|except| block.}
\label{lst:signals}
\end{listing}

Pausing actions is more complex (the action's internal state must be saved, a
means to resume the action must be provided and the user must be given ways to
code transition behaviours when pausing and resuming) and is not currently
supported by \pyRobots{}.

@@

Sharing the same syntax with {\sc urbi}, events in \pyRobots{} can by monitored
through the methods \python|on| (one shot event) and \python|whenever|
(continuous event monitoring). Event conditions can either be a predicate
(taking the robot instance as its only argument) or a $\{\text{state
key}, \text{condition}\}$ pair, with condition one of $\{\text{value}=x,
\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$.
Examples of event usage can be seen at lines 33--38 of
listing~\ref{croquignole_no_move}.

@@

\subsection{Resource management}

Physical robot resources (mainly actuators, but possibly sensors as
well) usually mandate exclusive use. \pyRobots{} provides a resource
locking mechanism that relies on the \python|@lock| annotation: an action
annotated with \python|@lock(ARMS)| (Listing~\ref{lst:resources}, l.6) will
acquire (or wait for) the \python|ARMS| resource before being executed.

This mechanism is generic and \pyRobots{} makes no assumptions concerning the
actual resources available to a specific robot. The
programmer define them himself, either as stand-alone or as compound
resources (Listing~\ref{lst:resources}, l.1--3).

A given action can lock several distinct resources with the corresponding
number of \python|@lock| annotations. When executed, the
action acquires a lock on the resource (and all sub-resources in the case
of a compound resource) which is released upon action completion. If the lock is
already acquired, two strategies are possible: either wait or \emph{fail fast}
(\python|@lock(..., wait=False)|, Listing~\ref{lst:resources}, l.16).

@@

Managing poses, transformations and frames in real-world scenarios is not an
easy task, mainly because it appears that libraries, middlewares or robotic
platforms tend each to rely on their own conventions; rotation representations,
in particular, are a common source of inconsistencies -- quaternions, Euler
conventions, rotation matrices, Rodrigues, yaw-pitch-roll notations being all
rather frequent. Within the ROS ecosystem, the Transform Library ({\sc
tf},~\cite{foote2013tf}) has played a major role in making it convenient to work
with frames: poses can always be expressed and shared in the most convenient
local frame, {\sc tf} taking care of required intermediate transformation.

\pyRobots{} pose management builds on this idea to offer flexible pose
representations \emph{within} the control scripts. Every time a 6D pose needs to
be specified (like a navigation goal or a target pose for an arm), it can be
expressed as a simple frame name, by a 3-tuple $(x, y, z)$ (default frame {\tt
map} is then assumed), by a 4-tuple $(x, y, z, frame)$, by a 6-tuple $(x, y, z,
rx, ry, rz)$ (using $sxyz$ Euler convention), by a 7-tuple $(x, y, z, rx, ry,
rz, frame)$, by a 7-tuple $(x, y, z, qx, qy, qz, qw)$, by an 8-tuple $(x, y, z,
qx, qy, qz, qw, frame)$ or by a dictionary that defines any subset of $(x, y, z,
qx, qy, qz, qw, frame)$.  Internally, \pyRobots{} normalizes and stores the
poses in their full 8-tuple expansion, as illustrated in
Listing~\ref{lst:poses}.

@@

Using helper methods such as \python|inframe| \python|(<pose>, <other frame>)| or
\python|pantilt| \python|(<pose>, <ref frame>)|, the programmer can then
transform a given pose into a different frame or different representation.

Required transformations are computed by \pyRobots{}, and the programmer can
provide new ``sources'' of frame transformations through the instantiation of
\python|FrameProviders| that takes advantage of robot- and middleware-specific
infrastructures (for example, ROS {\sc tf}, Aldebaran's {\tt NAOqi} frame
conventions, etc.).

@@

Running robots in real-world  environments that are dynamic and unpredictable
entails repetitive coding-testing-debugging development cycles. \pyRobots{}
provides several tools to support this development phase that rely on the
versatile Python built-in logging tools, which support configurable parallel
logging streams (typically, a full debug stream is logged to a file while a less
verbose stream is output on the console at run-time).

\pyRobots{} logs are designed to allow full \emph{offline} replay, and we
provide a graphical interface to display the actions and resource usage over
time (Figure~\ref{log_view}).

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{log}
        \caption{Screenshot of \pyRobots{} log analyzer, processing the log of the
        experiment presented in section~\ref{croq}}
        \label{log_view}
\end{figure}

Debugging and introspection support is provided at run-time through a range of
off-the-shelf Python debuggers. Remote debugging is also supported with standard
tools such as {\sc rpdb2}~\footnote{\url{http://winpdb.org}}. Python being an
interpreted language, these debuggers do not only provide insights on failures,
but also tooling for full introspection and run-time manipulation of the
application state.

@@

The development of \pyRobots{} started in 2011, driven by the requirements of a
18min long theatre performance between a human actor and a robot. Since that
time, the toolset has iteratively matured, and has been deployed and used on
several projects and platforms. We first briefly summarize these deployments,
and then present a recent study in a kindergarten that acted as a stress-test
for the library.

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
        \toprule
        Study                 & Robot & Middlewares \\ \midrule
        Roboscopie            & PR2       & ROS, {\sc GenoM} \\
        Interactive Grounding & PR2, LAAS-CNRS's Jido & ROS, {\sc GenoM} \\
        CoWriter              & Nao       & ROS, NAOqi  \\
        Ranger at kindergarten & EPFL's Ranger & ROS, aseba  \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the main studies where \pyRobots{} was used.}
\end{table}

\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
a theatre performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}.
The toolset was created out of the need to control a complex robot (a PR2) in a structured way, as required for a live performance with a
human\footnote{The original \pyRobots{} script used for the performance can be
accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
not reflect anymore the current toolkit usage.}.

A total of 47 \pyRobots{} actions were written for this event
(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolset
proved to be convenient for interleaving calls to both ROS-based and
{\sc GenoM}-based~\cite{mallet2010genom3} components through their respective
Python bindings.

\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience
of a script language, essential for the robot's programmer to quickly adapt to
the vision of the director.

@@

\subsubsection{Interactive Grounding} This study on perspective-taking and grounding
in human-robot interaction~\cite{lemaignan2013talking} is one of the several
experiments that was built on top of the LAAS-CNRS cognitive
architecture~\cite{lemaignan2014human}. In this large architecture, \pyRobots{}
is one of the two available executive controllers, and integrates with a task
planner, a symbolic knowledge base, several motion planners and a natural
language processor. \pyRobots{} was favoured over the second execution
controller (based on PRS) for smaller studies, as a more powerful alternative to
bare \textit{ad-hoc} scripts.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{cowriter}
        \caption{\textbf{CoWriter}: Children interactively teaching a Nao how to
        write.}
        \label{expe-cowriter}
\end{figure}

\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
(Figure~\ref{expe-cowriter}) involves a Nao robot and young students
in a \emph{learning by teaching} interaction: the children have to teach the
robot how to write. This experiment involved a dual ROS/NAOqi environment,
bridged within \pyRobots{}. The high-level behaviour was implemented as a finite
state machine ({\sc smach}) whose states corresponded to \pyRobots{} actions.
While the execution controller was subsequently simplified (towards a tighter
interaction scenario), this illustrates how \pyRobots{}, relying on the
ubiquitous Python language, can possibly blend into and interact with existing
tools.

@@

We organized a pilot study in a kindergarten to stress-test the hardware and
software of the Ranger robot~\cite{mondada2014ranger}. The Ranger robot is an
autonomous robot designed for interaction with children. As it was primarily
designed to support children with tidying up, it is shaped as a wooden ``box on
wheels'' (Figure~\ref{expe-nursery}). It features an RGB-D camera, IR sensors, a
range and bearing sensor, a physical front bumper, a scale, a removable pacifier
(used as a soft on/off switch by the children) and is covered with capacitive
tactile panels. The wooden panels hide 186 RGB LEDs, enabling various light
patterns, the two eyes can be fully controlled (2 DoFs pupils and actuated
eye-lids) and the robot can play sounds. Wheel controllers are designed such
that they switch to freewheel as soon as the robot is pushed by the children.
The robot is powered by a Gumstix Overo board (ARM7l) running a custom Linux
image, assisted by three custom microcontroller boards for low-level hardware
interfacing.

All user-facing behaviours are implemented with \pyRobots{},
and the communication with the hardware is abstracted through a mix of ROS
and Aseba~\cite{magnenat2011aseba} nodes.

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{\textbf{Stress-test}: Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviours of the robot are implemented via \pyRobots{}.}
        \label{expe-nursery}
\end{figure}
